---
title: "Example Analysis"
date: 2024-10-26
format: html
editor: visual
bibliography: references.bib
draft: false
---
**Question**
How do various U.S. commodities compare in their greenhouse gas emissions when assessed to the point of sale?

**Intended Audience**
This is a rough overview of the topic intended for the general public.

**Data Source**
This data comes from the U.S. EPA Office of Research and Development: https://catalog.data.gov/dataset/supply-chain-greenhouse-gas-emission-factors-v1-3-by-naics-6

**Data Dictionary**
https://github.com/USEPA/supply-chain-factors/blob/5df4328f9b06e69f7b6d21af0baf391de669974e/format-specs/README.md

**Data Wrangling**
```{r setup}
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
emissions <- read.csv('/Users/dhruthi11/Desktop/emissions.csv')
```

```{r rename}
emissions <- emissions %>% rename("Class Code" = X2017.NAICS.Code, "Commodity" = X2017.NAICS.Title, "Cradle to Production Emissions" = Supply.Chain.Emission.Factors.without.Margins, "Production to Sale Emissions" = Margins.of.Supply.Chain.Emission.Factors, "Cradle to Sale Emissions" = Supply.Chain.Emission.Factors.with.Margins, "Reference USEEIO Code" = Reference.USEEIO.Code)
```

```{r data}
em_arranged <- emissions %>% arrange(desc("Cradle to Sale Emissions"))

emissions2 <- emissions %>% select('Class Code', 'Commodity', "Cradle to Production Emissions", "Production to Sale Emissions", 'Cradle to Sale Emissions') 

#Top 3 Highest Cradle to Sale Emissions
emissions2 %>% slice_head(n=3)
#Top 3 Lowest Cradle to Sale Emissions
emissions2 %>% slice_tail(n=3)
```

```{r check values}
#Check if cradle to sale emissions are a combination of the other two emission categories (rounding could cause +/- 1 error)
check_values <- (abs(emissions2$'Cradle to Production Emissions' + emissions2$'Production to Sale Emissions') - emissions2$'Cradle to Sale Emissions') <= 1

if (all(check_values)) {
  result <- "All columns add up."
} else {
  result <- "NOT all columns add up."
}
print(result)
```

```{r plots}
#What is the distribution of Cradle to Sale Emissions?
emissions_long <- emissions2 %>%
  pivot_longer(cols = c(`Cradle to Production Emissions`, `Production to Sale Emissions`, `Cradle to Sale Emissions`), names_to = "Emission Type", values_to = "emissions")

ggplot(emissions_long, aes(emissions)) +
  geom_histogram(binwidth = 0.15) +
  facet_wrap(~ factor(`Emission Type`, 
                    levels = c("Cradle to Production Emissions", "Production to Sale Emissions", "Cradle to Sale Emissions"))) +
  labs(x = "Emissions", y = "Count") +
  ggtitle("Emission Levels per Production Step") +
  theme_minimal()
```

::: {.callout-note}
It appears that the Production to Sale emissions are generally less than Cradle to Production emissions - GHG emissions are released more during the cradle to production phase of making a commodity.
:::

```{r more plots}
#Is there a correlation between Cradle to Production Emission and Production to Sale Emission?
ggplot(emissions2, aes(x = `Cradle to Production Emissions`, y = `Production to Sale Emissions`)) +
  geom_point() +
  geom_smooth(method='lm') +
  ggtitle("Production to Sale versus Cradle to Production Emissions")
```

::: {.callout-note}
There seems to be a trend here where when a commodity releases more GHGs in the cradle to production phase, it is positively correlated with releasing more GHGs in the production to sale phase as well.
:::

```{r plots3}
#Which sector releases the most GHGs?
sector <- read.csv('/Users/dhruthi11/Desktop/class codes.csv')
emissions2$sectors <- str_extract(emissions2$`Class Code`, "\\d{2}")
emissions2$sectors <- as.integer(emissions2$sectors)
emissions3 <- emissions2 %>% full_join(sector, by = c('sectors' = 'Sector'))

#missing sector does not have information available
emissions3 %>% filter(is.na(Definition) | is.na(`Cradle to Sale Emissions`))

ggplot(emissions3, aes(factor(Definition), `Cradle to Sale Emissions`)) +
    geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle("Total Emissions of Production per Sector")
```

::: {.column-margin}
| Sector | Definition                                                               |
|--------|--------------------------------------------------------------------------|
| 11     | Agriculture, Forestry, Fishing and Hunting                               |
| 21     | Mining, Quarrying, and Oil and Gas Extraction                            |
| 22     | Utilities                                                                |
| 23     | Construction                                                             |
| 31 - 33     | Manufacturing                                                            |
| 42     | Wholesale Trade                                                          |
| 44 - 45    | Retail Trade                                                             |
| 48 - 49    | Transportation and Warehousing                                           |
| 51     | Information                                                              |
| 52     | Finance and Insurance                                                    |
| 53     | Real Estate and Rental and Leasing                                       |
| 54     | Professional, Scientific, and Technical Services                         |
| 55     | Management of Companies and Enterprises                                  |
| 56     | Administrative and Support and Waste Management and Remediation Services |
| 61     | Educational Services                                                     |
| 62     | Health Care and Social Assistance                                        |
| 71     | Arts, Entertainment, and Recreation                                      |
| 72     | Accommodation and Food Services                                          |
| 81     | Other Services (except Public Administration)                            |
| 92     | Public Administration  
:::

::: {.callout-note}
The Manufacturing sector releases much more GHGs than any other sector in the U.S. from cradle to sale.
:::

**Summary**

Data from the US EPA Office of R&D outlines emissions released from each sector of commodities manufactured in the US to compare the emissions per stage (Cradle to Production and Production to Sale) as well as between sector. Production and Sale Emissions are lower then Cradle to Production Emissions for most or all commodities, with a positive correlation between the emissions of the two stages. Manufacturing is a huge source of emissions with Culture, Forestry, Fishing, and Hunting, as well as Transportation and Warehousing being second and third highest in emissions. Public Administration and Management of Companies and Enterprises are the lowest in total emissions. These results can be used to determine which US commodities and production sectors need to be targetting for reducing emissions (manufacturing) through research and policy. It would be interesting to look into what sorts of research has been done using this dataset.

::: {.callout-important}
**Functions Used:**

- **dplyr:**
    - `rename`
    - `arrange`
    - `desc`
    - `select`
    - `slice_head`
    - `slice_tail`
    - `full_join`
    - `filter`
- **tidyr:**
    - `pivot_longer`
- **ggplot:**
    - `geom_histogram`
    - `geom_point`
    - `geom_smooth`
    - `geom_bar`
:::